\documentclass[]{article}
\voffset=-1.5cm
\oddsidemargin=0.0cm
\textwidth = 470pt
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{framed}

\usepackage{multicol}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{multicol}




%opening
\title{Binary Arithmetic - Tutorial Sheet}
\author{www.MathsResource.com}

\begin{document}
\section*{Part C: Number Bases - Hexadecimal}
\begin{enumerate}

\item Answer the following questions about the hexadecimal number systems
\begin{itemize}
\item[a)] How many characters are used in the hexadecimal system?
\item[b)] What is highest hexadecimal number that can be written with two characters? \item[c)] What is the equivalent number in decimal form?
\item[d)] What is the next highest hexadecimal number?
\end{itemize}


\item Which of the following are not valid hexadecimal numbers?
  \begin{multicols}{4}
    \begin{itemize}
    \item[a)] $73$
    \item[b)] $A5G$
    \item[c)] $11011$
    \item[d)] $EEF	$
    \end{itemize}
  \end{multicols}

\item Express the following decimal numbers as a hexadecimal number.
  \begin{multicols}{4}
    \begin{itemize}
    \item[a)] $(73)_{10}$
    \item[b)] $(15)_{10}$
    \item[c)] $(22)_{10}$
    \item[d)] $(121)_{10}$
    \end{itemize}
  \end{multicols}


\item Compute the following hexadecimal calculations.
  \begin{multicols}{4}
    \begin{itemize}
    \item[a)] $5D2+A30$
    \item[b)] $702+ABA$
    \item[c)] $101+111$
    \item[d)] $210+2A1$
    \end{itemize}
  \end{multicols}
\end{enumerate}



%-------------%-------------%-------------%----------------
\subsection*{Question 7}
\begin{itemize}
	

\item[(a)] What is highest Hexadecimal number that can be written with two characters, and what is it's equivalent in decimal form?
What is the next highest hexadecimal number?

% \[FF = 255\] % Remark 
\item[(b)] Which of the following are not valid hexadecimal numbers?

\begin{multicols}{2}
\begin{itemize}
\item[(i)] A5G 	
\item[(ii)] 73 
\item[(iii)] EEF	
\item[(iv)] 101
\end{itemize}
\end{multicols}
\end{itemize}	
%-------------%-------------%-------------%----------------
\subsection*{Question 10}

Calculate working in hexadecimal
\begin{itemize}
\item $(BBB)_{16} + (A56)_{16}$
\item $(BBB)_{16} - (A56)_{16}$
\end{itemize} 


%-------------%-------------%-------------%----------------
\subsection*{Question 11}

Write the hex number $(EC4)_{16}$ in binary.
Write the binary number $(11110110101|)_2$ in hex.
%-------------%-------------%-------------%----------------
%-------------%-------------%-------------%----------------
\subsection*{Question 12}

Express the decimal number 753 in binary , base 5 and hexadecimal.

\end{document}
